Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-07-09 thirdPaper/draft/typeConeAssociahedra.tex
coro:simplicialTypeConeCA typo
main results part II

Let~$\quiver = (Q,I)$ be a gentle bound quiver, and let~$\Fan(\quiver)$ be the fan of its~$\b{g}$-vectors (\cref{subsubsec:nonkissingComplex}), also called its non-kissing fan.
In \cref{sect::Kbproj}, we make use of representation theory in order to prove:

@@

@@

@@

\begin{corollary}
\label{coro:simplicialTypeConeCA}
The type cone~$\typeCone \big( \gvectorFan[n] \big)$ is simplicial.
\end{corollary}

>>>>

Let~$\quiver = (Q,I)$ be a gentle bound quiver, and let~$\Fan(\quiver)$ be the fan of its~$\b{g}$-vectors (\cref{subsubsec:nonkissingComplex}), also called its non-kissing fan.
We prove in \cref{sect::Kbproj}, by making use of representation theory:

@@

In \cref{sec:clusterCategories}, we prove an analogue for cluster categories of M. Auslander's result~\cite{Auslander1984}.
We are mainly motivated by the fact that itt has for a consequence that the type cones of the cluster fans of finite type are simplicial.
However, this result is of independent interest and sheds new lights on the Grothendieck groups of cluster categories.

Let~$\cat$ be a cluster category with finitely many isomorphism classes of indecomposable objects (see \cref{sect::setting} for the precise and more general setting in which the theorem is proven).
Let~$\Ksp(\cat)$ be the split Grothendieck group of~$\cat$.
Fix a cluster-tilting object~$T\in\cat$, and let~$K_0(\cat ; T)$ be the quotient of~$\Ksp(\cat)$ by the relations~$[X]+[Z]-[Y]$ for all triangles
$
X\xrightarrow{} Y \xrightarrow{} Z \xrightarrow{h} \susp X
$
with~$h \in (\susp T)$.
Denote by~$\b{g}:\Ksp(\cat) \to K_0(\cat ; T)$ the canonical projection.
For any indecomposable~$X\in\cat$, write~$\ell_X = [X]+[\susp^{-1}X]-[E]$, where~$E$ is the middle term of an almost split triangle starting at~$X$.
For any objects~$X,Y\in\cat$, we write~$\langle X, Y \rangle$ for~$\dim_\field\Hom{\Lambda}(FX,FY)$, where~$\Lambda$ is the cluster-tilted algebra~$\End{\cat}(T)$ and~$F$ is the equivalence of categories~$\cat(T,-):\cat/(\susp T) \to \MOD \Lambda$.

\noindent Theorem (see \cref{theo::relations-g-vecteurs}): \emph{The set~$L_\cat=\{\ell_X\,|\, X\in\ind(\cat)\setminus\add(\susp T)\}$ is a basis of the kernel of~$\b{g}$ and, for any~$x\in\ker(\b{g})$, we have:}
\[
x = \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]

@@

In \cref{sec:extricats} we prove a more general version of \cref{theo::relations-g-vecteurs} that not only applies to cluster fans, but also to non-kissing fans.
Once more, our main motivation is to prove that the type cone of the non-kissing fan of a brick and 2-acyclic gentle algebra is simplicial, but the results are of independent interest, and might be useful for studying the type cones of other families of simplicial fans.

Let~$\cat$ be an extriangulated category (\cref{sect::recollections extricats}) with a fixed projective object~$T$ such that the morphism~$T\to 0$ is an inflation and, for each~$X\in\cat$, there is an extriangle~$T_1^X\infl T_0^X\defl X\overset{\delta_X}{\dashrightarrow}$.
Let~$\susp T$ be the cone of the inflation~$T\infl 0$.

\noindent \cref{prop::KRextricat}: \emph{The functor~$\cat(T,-)$ induces an equivalence~$F:\cat/(\susp T) \to \MOD \End{\cat}(T)$.}

Our assumptions allow for a well-defined notion of index:~$\ind_T(X)=[T_0^X]-[T_1^X]\in\kzero{\add T}$.

\noindent \cref{prop::index iso extricat}: \emph{The index induces an isomorphism of abelian groups from~$\kzero{\cat}$ to~$\kzero{\add T}$.}

The previous two statements also hold when~$T$ is an additive subcategory rather than a mere object.
Assume moreover that~$\cat$ is~$\field$-linear, Ext-finite, Krull--Schmidt and has Auslander--Reiten--Serre duality.
Then, with notations analoguous to those used for cluster categories above, we obtain the following:

\noindent \cref{theorem::extricats}: \emph{The extriangulated category~$\cat$ has only finitely many ismorphism classes of indecomposable objects if and only if the set~$L_\cat$ generates the kernel of the canonical projection~$\b{g}:\Ksp(\cat)\to\kzero{\cat}$.
In that case~$L_\cat$ is a basis of the kernel of~$\b{g}$ and any~$x\in\ker(\b{g})$ decomposes as}
\[
 x = \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]

@@

\begin{corollary}
\label{coro:simplicialTypeConeCA}
The type cone~$\typeCone \big( \gvectorFan[\B_\circ] \big)$ is simplicial.
\end{corollary}
